---
title: "Covid91 vaccine study FinalFall2022"
author: "Put your name here!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

##Graphical Results 
```{r}
dd2 <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This barplot shows us that about 68% of people who took the placebo contracted COVID19 while about 32% of people who had the vaccine got COVID19. The vaccine seems to be more reliable than the placebo even tho some people on the vaccine caught covid. 

##Numerical 
```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```
This numerical result shows us that 68.44% of the placebo users caught COVID19 and 31.56% of vaccinated people caught COVID19 which means that a majority of vaccinated men are protected by the vaccination. 

##Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
Mfish <- fisher.test(table1)
Mfish
```
The inferential test shows us that there is a 95% confidence level and that interval is 1.732 for the placebo. 2.752 for the vaccines interval and the odds ratio is 2.178 the p value = 2.364e-12. 

```{r}
effect <- Mfish$p.value/(Mfish$p.value +1)
effectivenessM = Mfish$estimate/(Mfish$estimate +1)*100
print(effectivenessM)
```
The odds ratio is 16.53436 of the effectiveness on males. 
```{r}
low=(Mfish$conf.int[1]/(Mfish$conf.int[1] +1))*100
hi=(Mfish$conf.int[2]/(Mfish$conf.int[2] +1))*100
options(digits=2)
```

The high is 73.348 and the low is 63.402.
## Females


##Graphical
```{r}
dd2 <- F %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In this bar graph we can see that 58% of the women who took the vaccine got sick and 43% of women who took the placebo got sick. This means that the vaccine makes the women more likely to get sick. 

##Numerical




##Inferential




## LGBTQ



## Druggies


# Overall Results and Conclusions